---
title: "Test Rejection Sampler"
author: "Yihao Gu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source('rejection_sampler.R')
source('helper.R')
library('truncnorm')
library('truncdist')
```

## Case 1: $0 < a < 1, 0 \leq c < \frac{1}{a}$
```{r}
# Rejection sampling
a = 0.1
c = 5
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

## Case 2: $0 < a < 0.25, c \geq \frac{1}{a}$
```{r}
# Rejection sampling
a = 0.1
c = 20
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

## Case 3: $a \geq 0.25, c \geq \max(1,\frac{1}{a})$
```{r}
# Rejection sampling
a = 1
c = 1
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

## Case 4: $a \geq 1, 0.5 \leq c < 1$
```{r}
# Rejection sampling
a = 2
c = 0.8
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

## Case 5: $a \geq 1, 0 \leq c < 0.5$
```{r}
# Rejection sampling
a = 3
c = 0.3
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

## Case 6: $0 < a \leq 1, \max(-\frac{1}{a}, -\frac{1}{2a^2}) \geq c < 0$
```{r}
# Rejection sampling
a = 0.01
c = - 5
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

## Case 7: $a \geq 1$ or $c \leq - \frac{1}{2a^2}$
```{r}
# Rejection sampling
a = 1
c = -10
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

## Case 8: $0 < a < \frac{1}{2}, -\frac{1}{2a^2} \leq c < -\frac{1}{a}$
```{r}
# Rejection sampling
a = 0.1
c = -12
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

The generation of truncated normal random variables is a critical aspect of our process that warrants attention. For instance, consider the scenario where $a = 2, c = 0.3$ and we aim to generate a random variable from the truncated normal distribution over the interval $[0,0.3]$. Different R packages, specifically **truncnorm** and **truncdist**, exhibit varied performance in this task, which can significantly impact the results of our rejection sampler. 
```{r}
a = 2
c = 0.3
sd = sqrt(0.5) / a
sample1 = rtruncnorm(Nsim, a = 0, b = 0.3, mean = c, sd = sd)
sample2 = rtrunc(Nsim, spec = 'norm', a = 0, b = 0.3, mean = c, sd = sd)
hist(sample1, freq = FALSE)
hist(sample2, freq = FALSE)
```


