---
title: "Test Rejection Sampler"
author: "Yihao Gu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source('rejection_sampler.R')
source('helper.R')
library('truncdist')
library('truncnorm')
library('ggplot2')
library('dplyr')
library('purrr')
```

## Case 1: $0 < a < 1, 0 \leq c < \frac{1}{a}$
```{r}
# Rejection sampling
a = 0.99
c = 1 / a * 0.999
# c = 0
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

```{r}
# Check acceptance rate via simulation
aseq <- seq(1e-3, 0.99, length.out = 20)
c = seq(0, 0.99, length.out = 20)

# Create a dataframe
df <- map_df(aseq, ~{
  c_values = c / (.x)
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c_values, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="a * c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")
```


## Case 2: $0 < a < 0.25, c \geq \frac{1}{a}$
```{r}
# Rejection sampling
a = 0.001
c = 2 / a
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

```{r}
# Check acceptance rate via simulation
aseq <- seq(1e-3, 0.24, length.out = 20)
c = seq(1, 5, length.out = 20)

# Create a dataframe
df <- map_df(aseq, ~{
  c_values = seq(1 / (.x), 5 / (.x), length.out = 20)
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c_values, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="a * c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")
```


## Case 3: $a \geq 0.25, c \geq \max(1,\frac{1}{a})$
```{r}
# Rejection sampling
a = 2
c = 10
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

```{r}
# Check acceptance rate via simulation
aseq <- seq(0.25, 2, length.out = 20)
c = seq(1, 5, length.out = 20)

# Create a dataframe
df <- map_df(aseq, ~{
  start = max(1, 1/(.x))
  c_values = seq(start, 5 * start, length.out = 20)
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c_values, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="c/max(1, 1/a)", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")

```


## Case 4: $a \geq 1, 0.5 \leq c < 1$
```{r}
# Rejection sampling
a = 5
c = 0.99
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

```{r}
# Check acceptance rate via simulation
aseq <- seq(1, 5, length.out = 20)
c = seq(0.5, 0.99, length.out = 20)

# Create a dataframe
df <- map_df(aseq, ~{
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")

```


## Case 5: $a \geq 1, 0 \leq c < 0.5$
```{r}
# Rejection sampling
a = 10
c = 0.5 * 0.99
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

```{r}
# Check acceptance rate via simulation
aseq <- seq(1, 5, length.out = 20)
c = seq(0, 0.5 * 0.99, length.out = 20)

# Create a dataframe
df <- map_df(aseq, ~{
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")
```


## Case 6: $0 < a < 1, -\frac{1}{2a^2} \geq c < 0$
```{r}
# Rejection sampling
a = 0.99
c = - 0.5 / a^2
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```

```{r}
# Check acceptance rate via simulation
aseq <- seq(1e-3, 0.99, length.out = 20)
c = seq(1e-3, 1, length.out = 20)

# Create a dataframe
df <- map_df(aseq, ~{
  c_values = - c / 2 / (.x)^2
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c_values, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="-2a^2 * c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")
```

```{r}
# Check acceptance rate via simulation
aseq <- seq(1e-5, 1e-2, length.out = 20)
c = seq(1e-3, 1, length.out = 20)

# Create a dataframe
df3 <- map_df(aseq, ~{
  c_values = - c / 2 / (.x)^2
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c_values, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df3, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="-2a^2 * c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")
```


## Case 7: $a \geq 1$ or $c < - \frac{1}{2a^2}$
```{r}
# Rejection sampling
a = 0.99
c = - 0.5 / a^2 * 5
Nsim = 10^5
res = rejection_sampler(a, c, Nsim)
sample = res$sample
acceptance_rate = res$ar

# Acceptance rate
cat('The acceptance rate is', acceptance_rate, '.\n')

# Plot
plot_lscale_sample(sample, a, c)
```


When examining the acceptance rate, we divide regime 7 into two parts for convenience.
```{r}
# a >= 1, c < 0
# Check acceptance rate via simulation
aseq <- seq(1, 8, length.out = 20)
c = seq(-5,-1e-3, length.out = 20)

# Create a dataframe
df <- map_df(aseq, ~{
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")
```


```{r}
# a < 1, c < -0.5 / a^2
# Check acceptance rate via simulation
aseq <- seq(1e-3, 0.99, length.out = 20)
c = seq(1.01, 5, length.out = 20)

# Create a dataframe
df2 <- map_df(aseq, ~{
  start = - 0.5 / (.x)^2
  c_values = seq(1.01 * start, 5 * start, length.out = 20)
  data.frame(
    a = .x,
    c = c,
    acceptance_rate = sapply(c_values, get_acceptance_rate, a = .x)
  )
})

# Plot
ggplot(df2, aes(x=c, y=acceptance_rate, color=a, group=a)) + 
  geom_line() +
  labs(x="-2a^2 * c", y="Acceptance rate", color="a value") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_gradient(low="#72EDF2", high="#736EFE")
```

---
The generation of truncated normal random variables is a critical aspect of our process that warrants attention. For instance, consider the scenario where $a = 2, c = 0.3$ and we aim to generate a random variable from the truncated normal distribution over the interval $[0,0.3]$. Different R packages, specifically **truncnorm** and **truncdist**, exhibit varied performance in this task, which can significantly impact the results of our rejection sampler. 
```{r}
a = 2
c = 0.2
sd = sqrt(0.5) / a
Nsim = 10^5
breaks = seq(0, c, length.out = 20)
sample1 = rtruncnorm(Nsim, a = 0, b = c, mean = c, sd = sd) # truncnorm
sample2 = rtrunc(Nsim, spec = 'norm', a = 0, b = c, mean = c, sd = sd) # truncdist
hist(sample1, freq = FALSE, breaks = breaks)
hist(sample2, freq = FALSE, breaks = breaks)
```


```{r}
a = 0.1
c = 10
sd = sqrt(0.5) / a
Nsim = 10^5
breaks = seq(c/2, c, length.out = 20)
sample1 = rtruncnorm(Nsim, a = c/2, b = c, mean = c, sd = sd) # truncnorm
sample2 = rtrunc(Nsim, spec = 'norm', a = c/2, b = c, mean = c, sd = sd) # truncdist
hist(sample1, freq = FALSE, breaks = breaks)
hist(sample2, freq = FALSE, breaks = breaks)
```




